### 2011-12-16 article/Article.tex
mise a jour de la structure du document

%
% Complete documentation on the extended LaTeX markup used for Insight
% documentation is available in ``Documenting Insight'', which is part
% of the standard documentation for Insight.  It may be found online
% at:
%
%     http://www.itk.org/

\documentclass{InsightArticle}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}
% to be able to use options in graphics
\usepackage{graphicx}
% for pseudo code
\usepackage{listings}
% subfigures
\usepackage{subfigure}

%  This is a template for Papers to the Insight Journal. 
%  It is comparable to a technical report format.

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{Efficient N-Dimensional surface estimation using Crofton formula and run-length encoding}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
% \release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Ga\"etan Lehmann{$^1$} {\small{and}} David Legland{$^2$}}
\authoraddress{{$^1$}INRA, UMR 1198; ENVA; CNRS, FRE 2857, Biologie du D\'eveloppement et Reproduction, Jouy en Josas, F-78350, France.\\
{$^2$}INRA, XXX, France.}

\begin{document}
\maketitle

\ifhtml
\chapter*{Front Matter\label{front}}
\fi

\begin{abstract}
\noindent
Unlike the measure of the area in 2D or of the volume in 3D, the perimeter and the surface are not easily measurable in a discretized image.
In this article we describe a method based on the Crofton formula to measure those two parameters in a discritized image. The accuracy of
the method is discussed and tested on several known objects. An algorithm based on the run-length encoding of binary objects is presented
and compared to a brute force approach.
An implementation is provided and integrated in the LabelObject/LabelMap framework contributed earlier by the authors.
\end{abstract}

\tableofcontents

\section{Principles}
\subsection{Crofton formula}

\subsection{Implementation}

The number of intercepts, required to use the method described earlier, can be efficiently computed by using the run-length encoding of
the binary objects.

The implementation of this algorithm has been done in the \verb$itk::ShapeLabelMapFilter$ class, which is already in charge of the computation of
several shape descriptors. The run-length encoding used in the \verb$itk::LabelObject$ representation is reused.
The implementation is N-dimensional and thus is usable for any image dimension. The most useful cases, 2D and 3D have been specialized
to provide a more accurate estimation by also using the diagonals -- 4 directions in 2D and 13 directions in 3D. In the 4D case and greater,
the diagonals are not used.

\subsubsection{N-Dimensional name}
The attribute is called {\em Perimeter} independently of the dimension of the image and is available in the \verb$itk::ShapeLabelObject$.

\subsubsection{Border management}

The borders of the image is considered to be in the background, so the contour of an object touching the border of the image is measured.
This was not the case in the previous (undocumented) implementation and has been proven to be misleading for many users.
It is possible to subtract the perimeter on the border to the full perimeter if the measure without the part touching the border is needed.

\subsubsection{Multithreading}

The architecture implemented in \verb$itk::ShapeLabelMapFilter$ use one thread per LabelObject. The perimeter estimation has been integrated in this
architecture, providing a multithreaded implementation as long as there are several LabelObjects in the input LabelMap.

The perimeter estimation is now turned on by default, but can still disabled with \verb$SetComputePerimeter(false)$ in
\verb$itk::ShapeLabelMapFilter$ if not needed. This should enhance the user experience, especially for the attributes non obviously derived
from the perimeter like the roundness.

\subsection{Circularity}

\section{Evaluation method}
\subsection{Shape discretization}
\subsection{Test shapes}
\subsection{Measurments}
\subsection{Timing evaluation}

\section{Accuracy}
\subsection{Position effect}
\subsection{Orientation effect}

\section{Timing}

\section{Application}

ICOPAN

\appendix

\bibliographystyle{plain}
\bibliography{InsightJournal}
\nocite{ITKSoftwareGuide}

\end{document}

### 2011-12-16 article/Article.tex
update section names

@@
 
+\section{Introduction}
+
 \section{Principles}
@@
 
-\subsection{Implementation}
+\subsection{Intercepts count in digital images}
 
-The number of intercepts, required to use the method described earlier, can be efficiently computed by using the run-length encoding of
-the binary objects.
+The intercepts count, required to use the method described earlier, can be efficiently computed by using the run-length
+encoding of the binary objects.
 

### 2011-12-16 article/Article.tex
ajoute les equations de la partie maths

@@
 \authoraddress{{$^1$}INRA, UMR 1198; ENVA; CNRS, FRE 2857, Biologie du D\'eveloppement et Reproduction, Jouy en Josas, F-78350, France.\\
-{$^2$}INRA, XXX, France.}
+{$^2$}INRA, UMR 0782, G\'enie et Microbiologie des Proc\'ed\'es Alimentaire, Thiverval-Grignon, F-78850, France.}
 
@@
 \subsection{Crofton formula}
+
+% Crofton formula
+Perimeter measure of 2D objects, surface area measure of 3D objects 
+and more generally $(d-1)$-surface measure of $d$-dimensional objects 
+can be expressed in a unified formalism by using the Crofton formula.
+This formula consists in integrating the intercept number of the object with
+lines of various orientation and positions. Its expression in
+the general form is:
+\begin{eqnarray}
+S^{(d-1)}(X) & = & \frac{d v_d}{\pi}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
+\end{eqnarray}
+where $X$ is the structure of interest, $v_d$ is the volume of the $d$-dimensional ball,
+$\mathcal{L}^d$ is the set of all lines in the $d$-dimensional space. 
+The Euler-Poincar\'e characteristic $\chi$ is equal to number of connected components
+of the intersection of $X$ with a line $L$.
+
+% 2D and 3D
+For planar and 3D cases, Equation 1 can be rewritten:
+\begin{eqnarray}
+P(X) & = & \pi \int_{\mathcal{L}^2} \chi \left( X \cap L \right) dL \\
+S(X) & = &  4  \int_{\mathcal{L}^3} \chi \left( X \cap L \right) dL
+\end{eqnarray}
+
+% discretisation
+The Crofton formula can be easily applied to discrete binary images. 
+The integral over lines can be decomposed into an integral over a finite set of directions
+and an integral over lines parallel to a given direction.
 

### 2012-01-12 article/Article.tex
prcision sur le comptage des intersections
ajout de contenu (maths, methodo, precisions)

@@
 
+% to be able to use options in graphics
+\usepackage{graphicx}
+% for pseudo code
+\usepackage{listings}
+% subfigures
+\usepackage{subfigure}
+
+% some maths symbols
+\usepackage{amstext,amssymb}
+
+% to include accents in bliblio
+\usepackage[T1]{fontenc}
+\usepackage[latin1]{inputenc}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 %  hyperref should be the last package to be loaded.
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 ]{hyperref}
-% to be able to use options in graphics
-\usepackage{graphicx}
-% for pseudo code
-\usepackage{listings}
-% subfigures
-\usepackage{subfigure}
 
@@
 $\mathcal{L}^d$ is the set of all lines in the $d$-dimensional space. 
+The above integral is normalised such that the mass of lines hitting the unit ball 
+equals $v_{d-1}$, the measure of the unit ball projection on a $(d-1)$-dimensional plane.
+
 The Euler-Poincar\'e characteristic $\chi$ is equal to number of connected components
-of the intersection of $X$ with a line $L$.
+of the intersection of $X$ with a line $L$, or equivalently half the number of intersections
+of the boundary of $X$ with the line.
 
@@
 % discretisation
-The Crofton formula can be easily applied to discrete binary images. 
+The Crofton formula can be easily applied to discrete binary images \cite{Lang2001, Legland2007}. 
 The integral over lines can be decomposed into an integral over a finite set of directions
 and an integral over lines parallel to a given direction.
+The number of connected components in the intersections of the structure with a discrete line
+is obtained by counting the number of intercepts with its boundary.
 
@@
 
+The (planar) circularity (often called "shape factor") is defined as:
+\begin{eqnarray}
+f = 4\pi \frac{A}{p^2}
+\end{eqnarray}
+
+Circularity takes values between $0$ and $1$. It is equal to $1$ for circular shapes, and decreases for elongated shapes.
+
+[on garde la circularité ?]
+
 \section{Evaluation method}
+
+\subsection{Test shapes}
+
+Perimeter and surface area are measured on discretization of various 2D and 3D shapes 
+whose true perimeter or surface area is known.
+
+Planar test shapes include:
+\begin{itemize}
+\item disks
+\item ring, obtained by the difference of several disks
+\item trefoil shape
+\item ellipses with $a=XX$ and $b=XX$
+\item rectangle with length $XX$ and width $XX$
+\end{itemize}
+
+% % Example of reconstruction of a discrete disk
+% \begin{figure}[!htb]
+% \begin{center}
+% \includegraphics[width=5cm]{discreteDisk.eps}
+% %\includegraphics[width=5cm][bb=0 0 605 605]{chick.png}
+% \end{center}
+% \end{figure}
+
+Test shapes for 3D measurements include:
+\begin{itemize}
+\item balls with various radii
+\item hollow ball
+\item prolate and oblate ellipsoids
+\item cuboids
+\end{itemize}
+
 \subsection{Shape discretization}
-\subsection{Test shapes}
-\subsection{Measurments}
+
+A 2D or 3D binary image can be considered as a subset of a rectangular grid $L_d$ where
+d = 2,3. Such a grid can be written as
+\begin{eqnarray}
+L_d = \Delta_{1} \mathbb{Z} \times \ldots \times \Delta_{1} \mathbb{Z},
+\end{eqnarray}
+
+where the $d$-uple $(\Delta_{1},\ldots,\Delta_{d})$ defines the pixel
+or voxel size. A digitization scheme represents how a continuous
+structure $X$ is digitized into a binary image $X_\text{d}$. According to
+the lattice scheme, $X_\text{d} = X\cap\mathbb{L}^d$. According to the
+covering scheme a grid point $x$ belongs to $X_\text{d}$ if the grid
+cell centered on $x$ hits $X$. 
+This paper assumes isotropic resolution, i.e. $\Delta_{i}$ being the same for all $i$.
+
+\subsection{Measurements}
+
+For each test shape, measurements were made on
+several discretisation of the shape, with various orientations and various origins.
+
+The center of each shape was chosen at random within the center pixel or voxel of the grid.
+
 \subsection{Timing evaluation}
 
+???
+
 \section{Accuracy}
+
+\subsection{Global accuracy}
+
+[Tableaux qui donnent les valeurs moyennes pour les differentes formes]
+
+%
+\begin{table}[!htb]
+\begin{center}
+\begin{tabular}{|c|c|c|c|} \hline 
+shape		    & $U$		& $U_d^2$ & $U_d^4$		\tabularnewline \hline
+disk		    & $251.33$	& $251.31$ 	& $251.32$ 	\tabularnewline
+ring		    & $376.99$	& $376.99$ 	& $377.02$	\tabularnewline
+rectangle  $0$ 	& $120.00$ 	&\ \ $94.24$& $112.66$	\tabularnewline
+rectangle $30$  & $-$ 		& $127.23$	& $120.82$	\tabularnewline
+rectangle $45$  & $-$ 		& $131.95$	& $113.18$	\tabularnewline
+\hline
+\end{tabular}\end{center}
+\caption{\label{tab:Validate-Perimeter}
+Differences between theoretical and measured perimeter of planar shapes.
+Approximations were computed using $2$ and $4$ directions.
+}
+\end{table}
+
 \subsection{Position effect}
+
+[On regarde effet de la resolution et de la dispersion des mesures]
+[-> 1 graphe 2D, 1 graphe 3D, avec superposition des mesures ?]
+
 \subsection{Orientation effect}
@@
 \section{Timing}
+
+[c'est rapide]
 

### 2012-01-25 article/Article.tex
modifs mercredi 25/01

@@
 
+Dans l'intro, mettre :
+* mesure perimetre ou surface est important
+* liste des differentes methodes pour mesures surface / perimetre
+    * choix marching cubes pour comparer
+* parler des facteurs de forme et introduire "roundness"
+* plan de l'article
+
 \section{Principles}
@@
 
-% discretisation
+\subsection{Crofton formula in discrete images}
+
 The Crofton formula can be easily applied to discrete binary images \cite{Lang2001, Legland2007}. 
@@
 
+[Ajouter une formule avec des sommes a la place des integrales]
+
+[dire comment on calcule les poids (1) des espaces entre les droites et (2) des poids associes aux directions]
+[Cas des directions principales ($d$ en $nd$) et de toutes les directions (4 en 2D et 13 en 3D)]
+
 \subsection{Intercepts count in digital images}
@@
 encoding of the binary objects.
+
+\subsection{Roundness}
+
+The roundness is defined the ratio of equivalent $d$-perimeter over the measured $d$-perimeter:
+\begin{eqnarray}
+roundness = \frac{P_{eq}}{P}
+\end{eqnarray}
+
+Roundness takes values between $0$ and $1$. 
+It is equal to $1$ for circular or spherical shapes, 
+and decreases for more complicated shapes.
+
+\section{Implementation}
 
@@
 
-\subsubsection{N-Dimensional name}
+\subsection{N-Dimensional name}
 The attribute is called {\em Perimeter} independently of the dimension of the image and is available in the \verb$itk::ShapeLabelObject$.
 
-\subsubsection{Border management}
+\subsection{Border management}
 
@@
 
-\subsubsection{Multithreading}
+\subsection{Multithreading}
 
@@
 from the perimeter like the roundness.
-
-\subsection{Circularity}
-
-The (planar) circularity (often called "shape factor") is defined as:
-\begin{eqnarray}
-f = 4\pi \frac{A}{p^2}
-\end{eqnarray}
-
-Circularity takes values between $0$ and $1$. It is equal to $1$ for circular shapes, and decreases for elongated shapes.
-
-[on garde la circularité ?]
 
@@
 
-% % Example of reconstruction of a discrete disk
-% \begin{figure}[!htb]
-% \begin{center}
-% \includegraphics[width=5cm]{discreteDisk.eps}
-% %\includegraphics[width=5cm][bb=0 0 605 605]{chick.png}
-% \end{center}
-% \end{figure}
+ % Example of reconstruction of a discrete disk
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=5cm]{images/discreteDisk}
+\end{center}
+\end{figure}
 
@@
 The center of each shape was chosen at random within the center pixel or voxel of the grid.
+
+Perimeter of 2D discretized shapes was measured using Crofton method using 2 and 4 directions, 
+and by computing the length of the reconstructed contour (method ?).
+Surface area of 3D discretized shapes was measured using Crofton method using 3 and 13 directions, 
+and by computing the surface area of the reconstruced shape obtained by the marching cubes algorithm.
+The average and the standard deviation of measurements were computed for all shapes discretized
+with the same resolution. Effect of orientation was investigated by computing average and standard 
+deviation of shapes having the same orientation.
 

### 2012-01-25 article/Article.tex
eqn equivalent perimeter
ajoute equations et refs

@@
 
-The roundness is defined the ratio of equivalent $d$-perimeter over the measured $d$-perimeter:
+The roundness is defined the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
+The equivalent $d$-perimeter is defined as the perimeter of the $d$-ball with same volume $V$.
 \begin{eqnarray}
-roundness = \frac{P_{eq}}{P}
+roundness             & = & \frac{P_{eq}}{P} \\
+P_{eq}                & = & \frac{d \cdot V_{eq}}{R_{eq}} \\
+R_{eq}                & = & \sqrt[\displaystyle d]{\frac{V\cdot\Gamma(\frac{d+1}{2})}{\pi^\frac{d}{2}}} \\
+\Gamma(\frac{d+1}{2}) & = & \begin{cases}
+                               \displaystyle \frac{d}{2}!  &\text{if}~ d ~\text{is even} \\
+                               \displaystyle \frac{\sqrt{\pi} \cdot d!!}{2^{\frac{d+1}{2}}}  &\text{if}~ d ~\text{is odd} \\
+                            \end{cases} \\
+d!!                   & = & \begin{cases}
+                                1 &\text{if}~ d \leq 1 \\
+                                d \cdot (d-2)!!  & \text{otherwise} \\
+                             \end{cases}
 \end{eqnarray}
@@
 
-Planar test shapes include:
-\begin{itemize}
-\item disks
-\item ring, obtained by the difference of several disks
-\item trefoil shape
-\item ellipses with $a=XX$ and $b=XX$
-\item rectangle with length $XX$ and width $XX$
-\end{itemize}
+Planar test shapes include disks, rings obtained by the difference of several disks,
+trefoil shape, ellipses with $a=XX$ and $b=XX$, and rectangle with length $XX$ and width $XX$.
 
- % Example of reconstruction of a discrete disk
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics[width=5cm]{images/discreteDisk}
-\end{center}
-\end{figure}
-
-Test shapes for 3D measurements include:
-\begin{itemize}
-\item balls with various radii
-\item hollow ball
-\item prolate and oblate ellipsoids
-\item cuboids
-\end{itemize}
+Test shapes for 3D measurements include
+balls with various radii, hollow ball, prolate and oblate ellipsoids, and cuboids. 
+For each type of shape, several orientations were considered.
 
@@
 or voxel size. A digitization scheme represents how a continuous
-structure $X$ is digitized into a binary image $X_\text{d}$. According to
-the lattice scheme, $X_\text{d} = X\cap\mathbb{L}^d$. According to the
+structure $X$ is digitized into a binary image $X_\text{d}$. 
+According to the
 covering scheme a grid point $x$ belongs to $X_\text{d}$ if the grid
-cell centered on $x$ hits $X$. 
+cell centered on $x$ hits $X$ (See Fig.~\ref{fig:DiskDiscretization}).
+
+% Example of reconstruction of a discrete disk
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=5cm]{images/discreteDisk}
+\end{center}
+\caption{Discretization of a disk}
+\label{fig:DiskDiscretization}
+\end{figure}
 This paper assumes isotropic resolution, i.e. $\Delta_{i}$ being the same for all $i$.

### 2012-01-25 article/Article.tex
refonte biblio, MAJ des affiliations, et ajout de figures

@@
 \documentclass{InsightArticle}
+
+\usepackage{color}
 
@@
 
+ \definecolor{lightgray}{gray}{.75}
+\newcommand{\commentaire}[1]{
+\colorbox{lightgray}{#1}
+}
+
 %  This is a template for Papers to the Insight Journal. 
@@
 % snail-mail address if you like.
-\author{Ga\"etan Lehmann{$^1$} {\small{and}} David Legland{$^2$}}
-\authoraddress{{$^1$}INRA, UMR 1198; ENVA; CNRS, FRE 2857, Biologie du D\'eveloppement et Reproduction, Jouy en Josas, F-78350, France.\\
-{$^2$}INRA, UMR 0782, G\'enie et Microbiologie des Proc\'ed\'es Alimentaire, Thiverval-Grignon, F-78850, France.}
+\author{Ga\"etan Lehmann$^{1,2}$ {\small{and}} David Legland$^{3,4,5}$}
+
+\authoraddress{\small{
+{$^1$}INRA, UMR 1198 Biologie du D\'eveloppement et Reproduction, Jouy-en-Josas, F-78350, France\\
+{$^2$}ENVA, UMR 1198 Biologie du D\'eveloppement et Reproduction, Jouy-en-Josas, F-78350, France\\
+{$^3$}INRA, UMR 0782 G\'enie et Microbiologie des Proc\'ed\'es Alimentaire, Thiverval-Grignon, F-78850, France\\
+{$^4$}AgroParisTech, UMR 0782 G\'enie et Microbiologie des Proc\'ed\'es Alimentaire, Thiverval-Grignon, F-78850, France\\
+{$^5$}INRA, UMR 1318 Institut Jean-Pierre Bourgin, Versailles, F-78026, France
+}}
 
@@
 
-Dans l'intro, mettre :
-* mesure perimetre ou surface est important
-* liste des differentes methodes pour mesures surface / perimetre
-    * choix marching cubes pour comparer
-* parler des facteurs de forme et introduire "roundness"
-* plan de l'article
+Surface area and perimeter are widely used parameters describing the size of objects observed in images, 
+and are commonly used for computing various shape factors.
+Unlike the area that can be easily measured by counting pixels belonging to the object, measuring the 
+perimeter is not as straightforward, 
+and naive methods can lead to huge systematic errors \cite{Klette2004, Legland2007}.
+Usually, the countour of the object is extracted, then its length is measured.
+The same approach is usually applied to 3D images: the boundary of the 3D object is first reconstructed,
+e.g. by using marching cubes \cite{Lorensen1987}, then its surface area is computed by summing 
+area of individual triangles.
+
+The Crofton method is an alternative method that allows measuring the perimeter of 2D objects,
+the surface area of 3D objects, and more generally the $(d-1)$-dimensional measure of $d$-dimensional 
+object boundary. 
+\commentaire{Ajouter un petit argumentaire sur convergence des mesures avec Crofton.}
+
+The article first recalls the mathematical principles of $d$-perimeter measures in digital images
+using the Crofton method. An efficient implementation based on run-length encoding is then presented.
+The method is evaluated on various $2D$ and $3D$ synthetic shapes and compared with other methods.
+The effect on the roundness shape factor is also investigated.
 
@@
 \begin{eqnarray}
-S^{(d-1)}(X) & = & \frac{d v_d}{\pi}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
+S^{(d-1)}(X) & = & \frac{d \cdot v_d}{\pi}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
 \end{eqnarray}
@@
 is obtained by counting the number of intercepts with its boundary.
+\commentaire{ 
+Ajouter une formule avec des sommes a la place des integrales (David)}
 
-[Ajouter une formule avec des sommes a la place des integrales]
+\commentaire{dire comment on calcule les poids (1) des espaces entre les droites et (2) des poids associes aux directions}
 
-[dire comment on calcule les poids (1) des espaces entre les droites et (2) des poids associes aux directions]
-[Cas des directions principales ($d$ en $nd$) et de toutes les directions (4 en 2D et 13 en 3D)]
+\commentaire{Cas des directions principales ($d$ en $nd$) et de toutes les directions (4 en 2D et 13 en 3D)}
 
@@
 \end{figure}
-This paper assumes isotropic resolution, i.e. $\Delta_{i}$ being the same for all $i$.
 
@@
 Perimeter of 2D discretized shapes was measured using Crofton method using 2 and 4 directions, 
-and by computing the length of the reconstructed contour (method ?).
+and by computing the length of the reconstructed contour (Matlab Software was used).
 Surface area of 3D discretized shapes was measured using Crofton method using 3 and 13 directions, 
@@
 
-\subsection{Global accuracy}
+\subsection{Perimeter of disks at various resolutions}
 
-[Tableaux qui donnent les valeurs moyennes pour les differentes formes]
+Figure \ref{fig:MeasureDiskPerimeter} shows average and standard deviation of perimeter
+measured on disks with various resolutions. Perimeter measured using Crofton method converges
+towards the real value when the resolution increase. 
+The convergence speed is faster when four directions are used instead of two, 
+and the variability is lower.
 
-%
-\begin{table}[!htb]
+% Example of reconstruction of a discrete disk
+\begin{figure}[!htb]
 \begin{center}
-\begin{tabular}{|c|c|c|c|} \hline 
-shape		    & $U$		& $U_d^2$ & $U_d^4$		\tabularnewline \hline
-disk		    & $251.33$	& $251.31$ 	& $251.32$ 	\tabularnewline
-ring		    & $376.99$	& $376.99$ 	& $377.02$	\tabularnewline
-rectangle  $0$ 	& $120.00$ 	&\ \ $94.24$& $112.66$	\tabularnewline
-rectangle $30$  & $-$ 		& $127.23$	& $120.82$	\tabularnewline
-rectangle $45$  & $-$ 		& $131.95$	& $113.18$	\tabularnewline
-\hline
-\end{tabular}\end{center}
-\caption{\label{tab:Validate-Perimeter}
-Differences between theoretical and measured perimeter of planar shapes.
-Approximations were computed using $2$ and $4$ directions.
-}
-\end{table}
+\includegraphics[width=10cm]{images/perimeterDisks}
+\end{center}
+\caption{Perimeter measure of disks with various resolutions and at random positions}
+\label{fig:MeasureDiskPerimeter}
+\end{figure}
 
-\subsection{Position effect}
+Perimeters measured with Matlab show similar variability, but the measured values do not
+converge toward the true value. This shows the inaccuracy of the method consisting in
+counting boundary pixels.
 
-[On regarde effet de la resolution et de la dispersion des mesures]
-[-> 1 graphe 2D, 1 graphe 3D, avec superposition des mesures ?]
+The oscillations of the variability is due to the fact that for disks with diameters equal to an
+integer, the number of intersections with horizontal or vertical lines is always the same, 
+thus greatly reducing the variability of measures.
 
-\subsection{Orientation effect}
+\subsection{Perimeter of squares with various orientations}
+
+Figure \ref{fig:MeasureSquareOriented} shows measures obtained on discretized squares (side 30 pixels)
+with various orientations. All methods oscillate around the expected value. 
+Using Crofton method with only 2 directions can produce greater errors on the measure than with 4 
+directions.
+
+% Example of reconstruction of a discrete squares with various orientations
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=8cm]{images/errorSquareRotated}
+\includegraphics[width=8cm]{images/errorRelSquareRotated}
+\end{center}
+\caption{Variations of perimeter measure of a digitized square depending on the orientation}
+\label{fig:MeasureSquareOriented}
+\end{figure}
 
@@
 
-ICOPAN
+\subsection{ICOPAN}
+
+Methods were applied on images acquired within the ICOPAN project \cite{Andrey2010}.
 

### 2012-02-03 article/Article.tex
complete the intercept count algorithm

@@
 ]{hyperref}
+% to be able to use options in graphics
+\usepackage{graphicx}
+% for pseudo code
+\usepackage{listings}
+% subfigures
+\usepackage{subfigure}
+\usepackage{pseudocode}
 
@@
 
+The run-length encoding groups all the pixels on the same line in a single element.
+If the binary object is compact enough, as it is often the case for connected components, the run-length encoding heavily reduces the number of elements required to represent the object, compared to the simple pixel representation.
+Thanks to the run-length encoding, it is possible to compute the intercepts the binary object line by line, instead of computing it pixel by pixel.
+
+The data structure used as the input of the algorigthm is a set of all the indexes of the object with their first element remove -- in 3D, this is all the possible (y, z) pairs -- to which are associated a list of lines.
+The lines are coded as a $x$ position and a length.
+
+The algorithm used to count the intercepts is described in Figure \ref{interceptCountAlgo}.
+
+\begin{figure}[htbp]
+\centering
+\small
+\begin{pseudocode}[framebox]{countIntercept}{O}
+\COMMENT{intercept count} \\
+ic \GETS \emptyset \\
+\COMMENT{the offset of the neighbors on the first axis} \\
+xno \GETS offset(0) \\
+xno[0] \GETS 1 \\
+\FOREACH ls \in O \DO
+\BEGIN
+  \COMMENT{2 intercepts per line on the first axis} \\
+  ic[xno] \GETS ic[xno] + 2 \cdot length(ls) \\
+  \FOREACH nls \in neighbors(ls) \DO
+  \BEGIN
+    \COMMENT{straight and diagonal offsets of the neighbors} \\
+    no \GETS offset(nls) \\
+    dno \GETS offset(nls) \\
+    dno[0] \GETS 1 \\
+    \IF isEmpty(nls) \THEN
+    \BEGIN
+      \COMMENT{all the lines are on a contour} \\
+      \FOREACH l \in ls \DO
+      \BEGIN
+	ic[no] \GETS ic[no] + length(l)\\
+	ic[dno] \GETS ic[dno] + 2 \cdot length(l)
+      \END
+    \END
+    \ELSE
+    \BEGIN
+      \COMMENT{iterate over the lines in $ls$ and the empty lines in $nls$} \\
+      il \GETS iterator(ls) \\
+      inl \GETS iterator(nls) \\
+      nMin \GETS -\infty \\
+      nMax \GETS pos(inl) - 1 \\
+      \WHILE isNotAtEnd(il) \DO
+      \BEGIN
+        \COMMENT{measure the intersection of these two lines} \\
+	ic[no] \GETS ic[no] + max( 0, min( lMax, nMax ) - max( lMin, nMin ) + 1 ) \\
+	ic[dno] \GETS ic[dno] + max( 0, min( lMax, nMax+1 ) - max( lMin, nMin+1 ) + 1 )\\
+	ic[dno] \GETS ic[dno] + max( 0, min( lMax, nMax-1 ) - max( lMin, nMin-1 ) + 1 )\\
+        \COMMENT{and move to the next line, either in $ls$ or in $nls$} \\
+	\IF nMax \leq lMax \THEN
+	\BEGIN
+          nMin \GETS pos(inl) + length(inl) \\
+          next(inl) \\
+          \IF isNotAtEnd(inl) \DO
+            nMax \GETS pos(inl) - 1
+          \ELSE
+            nMax \GETS \infty
+	\END
+	\ELSE
+          next(il)
+      \END
+    \END
+  \END
+\END
+\end{pseudocode}
+\caption{\label{interceptCountAlgo}The algorithm used to count the intercepts}
+\end{figure}
+
 \subsection{Roundness}
@@
 
+The perimeter estimation is now turned on by default, but can still be disabled with \verb$SetComputePerimeter(false)$ in
+\verb$itk::ShapeLabelMapFilter$ if not needed. This should enhance the user experience, especially for the attributes non obviously derived
+from the perimeter like the roundness.
+
 \subsection{N-Dimensional name}
@@
 architecture, providing a multithreaded implementation as long as there are several LabelObjects in the input LabelMap.
-
-The perimeter estimation is now turned on by default, but can still disabled with \verb$SetComputePerimeter(false)$ in
-\verb$itk::ShapeLabelMapFilter$ if not needed. This should enhance the user experience, especially for the attributes non obviously derived
-from the perimeter like the roundness.
 

### 2012-02-06 article/Article.tex
add missing computation of lMin and lMax
ajoute equations sommes, ajoute schema poids des directions, biblio

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%  hyperref should be the last package to be loaded.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[dvips,
-bookmarks,
-bookmarksopen,
-backref,
-colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
-]{hyperref}
 % to be able to use options in graphics
@@
 }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  hyperref should be the last package to be loaded.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[
+bookmarks,
+bookmarksopen,
+backref,
+colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
+]{hyperref}
 
@@
 
-The Crofton method is an alternative method that allows measuring the perimeter of 2D objects,
+The Crofton method is an alternative method that allows estimating the perimeter of 2D objects,
 the surface area of 3D objects, and more generally the $(d-1)$-dimensional measure of $d$-dimensional 
 object boundary. 
-\commentaire{Ajouter un petit argumentaire sur convergence des mesures avec Crofton.}
+It is based on counting intercept number of the object boundary with a set of isotropic test lines, 
+and provide an unbiased estimate of the actual perimeter or surface.
 
@@
 \begin{eqnarray}
-S^{(d-1)}(X) & = & \frac{d \cdot v_d}{\pi}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
+S^{(d-1)}(X) & = & \frac{d \cdot v_d}{v_{d-1}}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
 \end{eqnarray}
@@
 The integral over lines can be decomposed into an integral over a finite set of directions
-and an integral over lines parallel to a given direction.
-The number of connected components in the intersections of the structure with a discrete line
-is obtained by counting the number of intercepts with its boundary.
-\commentaire{ 
-Ajouter une formule avec des sommes a la place des integrales (David)}
+and an integral over all the lines parallel to a given direction.
 
-\commentaire{dire comment on calcule les poids (1) des espaces entre les droites et (2) des poids associes aux directions}
+For the planar case, the perimeter can be estimated by considering horizontal and vertical lines, 
+i.e.\ two directions. An alternative is to use the diagonals, resulting in four directions. 
+Perimeter estimate is then written as:
+\begin{eqnarray}
+P(X) & \simeq & \pi \sum_{k}\frac{c_{k}}{\lambda_{k}}\chi \left( X \cap L_{k} \right)
+\end{eqnarray}
+where $c_k$ is the discretization weight associated to direction $k$, 
+$\lambda_k$ is the density of discrete lines in direction $k$, 
+and $L_k$ is the set of all discrete lines in direction $k$.
 
-\commentaire{Cas des directions principales ($d$ en $nd$) et de toutes les directions (4 en 2D et 13 en 3D)}
+The line density $\lambda_k$ may vary according to the directions, 
+due to image resolution or to the use of diagonals. 
+It is computed as the ratio of the distance between two neighbor pixels with
+the area associated to a pixel.
+When only horizontal and vertical directions are used, associated weights $c_k$ equal $1/2$.
+When diagonals are used, weights are obtained by projecting direction vectors on the unit circle, 
+and by computing the relative fraction of circle associated to each direction 
+(Fig.~\ref{fig:DirectionWeights2D3D}).
+
+% Compute direction weights
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=8cm]{images/directionsWeights2D}
+\end{center}
+\caption{Computation of direction weights for rectangular pixel grid. 
+Black dots correspond to pixel centers. 
+Angular regions are computed between direction bisectors.}
+\label{fig:DirectionWeights2D3D}
+\end{figure}
+
+In a similar way, surface area of $3D$ objects may be estimated using following approximation:
+\begin{eqnarray}
+S(X) & \simeq & 4\sum_{k}\frac{c_{k}}{\lambda_{k}}\chi \left( X \cap L_{k} \right)
+\end{eqnarray}
+where $L_k$ is the set of $3D$ discrete lines parallel to direction $k$, 
+$c_k$ the discretization weight associated to $3D$ direction $k$, 
+and $\lambda_k$ the density of discrete lines in direction $k$.
+
+Line densities are computed as for the planar case, by dividing distance of neighbor
+voxels in direction $k$ by the volume associated to a single voxel.
+When surface area is estimated by taking into account the three main direction in images, 
+associated weights $c_k$ equal $1/3$. If more directions are considered, the strategy of
+\cite{Ohser2000} was used: each direction was projected on the unit sphere, 
+and used as a germ for computing Voronoï diagram on the unit sphere. 
+The relative surface area of each spherical domain was used as weight for the
+corresponding direction.
 
@@
 
-The intercepts count, required to use the method described earlier, can be efficiently computed by using the run-length
+The number of connected components in the intersections of the structure with a discrete line
+is obtained by counting the number of intercepts with its boundary.
+This intercepts count can be efficiently computed by using the run-length
 encoding of the binary objects.
@@
       \BEGIN
+        lMin \GETS pos(il) \\
+        lMax \GETS pos(il) + length(il) - 1 \\
         \COMMENT{measure the intersection of these two lines} \\
@@
 
-The roundness is defined the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
+The roundness is defined as the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
 The equivalent $d$-perimeter is defined as the perimeter of the $d$-ball with same volume $V$.
@@
 
-\subsection{Shape discretization}
-
-A 2D or 3D binary image can be considered as a subset of a rectangular grid $L_d$ where
+Shapes were discretized following the Gauss discretization scheme \cite{Klette2004, Legland2007}.
+Binary images can be considered as a subset of a rectangular grid $L_d$ where
 d = 2,3. Such a grid can be written as
@@
 \end{eqnarray}
-
 where the $d$-uple $(\Delta_{1},\ldots,\Delta_{d})$ defines the pixel
-or voxel size. A digitization scheme represents how a continuous
-structure $X$ is digitized into a binary image $X_\text{d}$. 
-According to the
-covering scheme a grid point $x$ belongs to $X_\text{d}$ if the grid
+or voxel size. 
+A grid point $x$ belongs to the reconstructed structure if the grid
 cell centered on $x$ hits $X$ (See Fig.~\ref{fig:DiskDiscretization}).

### 2012-02-10 article/Article.tex
quelques images de plus, et un tableau de resultats

@@
 \begin{center}
-\includegraphics[width=8cm]{images/directionsWeights2D}
+\includegraphics[width=8cm]{images/directionWeights2D}
+\includegraphics[width=7cm]{images/directionWeights3D_335}
 \end{center}
-\caption{Computation of direction weights for rectangular pixel grid. 
-Black dots correspond to pixel centers. 
-Angular regions are computed between direction bisectors.}
+\caption{Computation of direction weights for rectangular grids. 
+Left: planar case. Black dots correspond to pixel centers. 
+Angular regions are computed between direction bisectors.
+Right: 3D case. The unit sphere is partitioned into spherical polygons
+corresponding to each of the 26 oriented directions.
+}
 \label{fig:DirectionWeights2D3D}
@@
 associated weights $c_k$ equal $1/3$. If more directions are considered, the strategy of
-\cite{Ohser2000} was used: each direction was projected on the unit sphere, 
-and used as a germ for computing Voronoï diagram on the unit sphere. 
+Ohser and M\"ucklich \cite{Ohser2000} was applied: each direction was projected on the unit sphere, 
+and was used as a germ for computing Voronoï diagram on the unit sphere. 
 The relative surface area of each spherical domain was used as weight for the
-corresponding direction.
+corresponding direction (Fig.~\ref{fig:DirectionWeights2D3D}).
 
@@
 \begin{center}
-\includegraphics[width=5cm]{images/discreteDisk}
+\includegraphics[width=12cm]{images/discreteDisks2}
 \end{center}
-\caption{Discretization of a disk}
+\caption{Discretization of two disks with same radius and different positions. 
+A grid cell belong tho the reconstructed shape if its center is inside the original shape.
+The reconstructed shapes have different size and shape.}
 \label{fig:DiskDiscretization}
@@
 several discretisation of the shape, with various orientations and various origins.
-
 The center of each shape was chosen at random within the center pixel or voxel of the grid.
@@
 
-\subsection{Timing evaluation}
-
-???
+% \subsection{Timing evaluation}
+% ???
 
@@
 Using Crofton method with only 2 directions can produce greater errors on the measure than with 4 
-directions.
+directions. However, in both cases, the average over the orientations of the mean values is very close 
+to the actual perimeter.
 
@@
 \begin{center}
-\includegraphics[width=8cm]{images/errorSquareRotated}
-\includegraphics[width=8cm]{images/errorRelSquareRotated}
+\includegraphics[width=8cm]{images/perimRotatedSquares}
+\includegraphics[width=8cm]{images/errorRotatedSquares}
 \end{center}
-\caption{Variations of perimeter measure of a digitized square depending on the orientation}
+\caption{Variations of perimeter measure of a digitized square depending on the orientation. 
+Left: perimeter measures. Right: relative errors (in percent).}
 \label{fig:MeasureSquareOriented}
@@
 
-\section{Timing}
+% TODO mesure pour des formes 3D
+% able
 
-[c'est rapide]
+\subsection{Surface area of 3D shapes}
+
+Comarison of surface area measured on several discrete reconstructed shapes is given in
+Table \ref{tab:CompareSurfaceArea}.
+As for the planar case, the measure is closer to the actual value when using more
+directions. The error is around 2 percent in average for 3 directions, and less than
+1 percent for 13 directions.
+
+When using isosurface reconstruction, the surface area is systematically over estimated.
+The difference is around 8 percent in average.
+
+\begin{table}[!htb]
+\begin{center}\begin{tabular}{ccccccc}
+shape	& $(\varphi,\theta)$ & $S$ 	& $S_{crofton}^3$ & $S_{crofton}^{13}$ & $S_{isosurface}$ \tabularnewline \hline 
+ball 		& $n.a.$  		& 	$11309.8$	& $11312.0$ 	& $11306.6$ 	& $12253.9$ 	\tabularnewline 
+prolate	& $( 0,  0)$  	& 	$3082.9$ 	& $2938.7$ 	& $3084.6$ 	& $3344.4$ 	\tabularnewline
+- 			& $(45,  0)$  	& 	$-$ 		& $3137.3$ 	& $3086.3$ 	& $3340.6$ 	\tabularnewline
+- 			& $(45, 45)$  	& 	$-$ 		& $3150.6$ 	& $3083.3$ 	& $3337.8$ 	\tabularnewline
+oblate 	& $( 0,  0)$  	& 	$6856.8$ 	& $6290.7$ 	& $6807.6$ 	& $7387.4$ 	\tabularnewline
+- 			& $(45,  0)$  	& 	$-$ 		& $6872.0$ 	& $6789.0$ 	& $7346.2$ 	\tabularnewline
+- 			& $(45, 45)$  	& 	$-$ 		& $7154.7$ 	& $6808.5$ 	& $7366.1$ 	\tabularnewline
+torus 		& $( 0,  0)$  	& 	$11843.5$	& $11417.3$ 	& $11792.9$ 	& $12785.4$ 	\tabularnewline
+-			& $(45,  0)$  	& 	$-$    	& $11809.3$ 	& $11766.8$ 	& $12844.4$ 	\tabularnewline
+- 			& $(45, 45)$ 	& 	$-$ 	 	& $12086.0$ 	& $11822.4$ 	& $12798.4$ 	\tabularnewline
+\end{tabular}\end{center}
+\caption{ \label{tab:CompareSurfaceArea}
+Differences between actual surface area and its measures with different methods, 
+on shapes with various orientations.
+The orientation is given by the direction of the shape rotation axis, defined by 
+the azimut $\varphi$ (between 0 and 360 degrees) and the colatitude $\theta$ (between 0 and 180 degrees).}
+\end{table}
+
+% \section{Timing}
+% [c'est rapide]
 

### 2012-02-17 article/Article.tex
quelques typos dans le doc

@@
 % snail-mail address if you like.
-\author{Ga\"etan Lehmann$^{1,2}$ {\small{and}} David Legland$^{3,4,5}$}
+\author{Ga\"etan Lehmann$^{1,2}$ {\small{and}} David Legland$^{3,4}$}
 
@@
 {$^4$}AgroParisTech, UMR 0782 G\'enie et Microbiologie des Proc\'ed\'es Alimentaire, Thiverval-Grignon, F-78850, France\\
-{$^5$}INRA, UMR 1318 Institut Jean-Pierre Bourgin, Versailles, F-78026, France
 }}
@@
 \section{Principles}
-\subsection{Crofton formula}
+\subsection{Surface area and perimeter estimation}
 
@@
 \begin{center}\begin{tabular}{ccccccc}
-shape	& $(\varphi,\theta)$ & $S$ 	& $S_{crofton}^3$ & $S_{crofton}^{13}$ & $S_{isosurface}$ \tabularnewline \hline 
+shape	& $(\varphi,\theta)$ & $S$ 	& $S_{crofton^3}$ & $S_{crofton^{13}}$ & $S_{isosurface}$ \tabularnewline \hline 
 ball 		& $n.a.$  		& 	$11309.8$	& $11312.0$ 	& $11306.6$ 	& $12253.9$ 	\tabularnewline 

### 2012-02-17 article/Article.tex
article continuated

@@
 }
+
+\let\oldtabular=\tabular
+\def\tabular{\small\oldtabular}
 
@@
 \begin{center}
-\includegraphics[width=8cm]{images/perimRotatedSquares}
-\includegraphics[width=8cm]{images/errorRotatedSquares}
+\includegraphics[width=8cm]{images/perimRotatedEllipses}
+\includegraphics[width=8cm]{images/errorRotatedEllipses}
 \end{center}
-\caption{Variations of perimeter measure of a digitized square depending on the orientation. 
+\caption{Variations of perimeter measure of a digitized ellipse depending on the orientation. 
 Left: perimeter measures. Right: relative errors (in percent).}
@@
 \begin{center}\begin{tabular}{ccccccc}
-shape	& $(\varphi,\theta)$ & $S$ 	& $S_{crofton^3}$ & $S_{crofton^{13}}$ & $S_{isosurface}$ \tabularnewline \hline 
-ball 		& $n.a.$  		& 	$11309.8$	& $11312.0$ 	& $11306.6$ 	& $12253.9$ 	\tabularnewline 
-prolate	& $( 0,  0)$  	& 	$3082.9$ 	& $2938.7$ 	& $3084.6$ 	& $3344.4$ 	\tabularnewline
-- 			& $(45,  0)$  	& 	$-$ 		& $3137.3$ 	& $3086.3$ 	& $3340.6$ 	\tabularnewline
-- 			& $(45, 45)$  	& 	$-$ 		& $3150.6$ 	& $3083.3$ 	& $3337.8$ 	\tabularnewline
-oblate 	& $( 0,  0)$  	& 	$6856.8$ 	& $6290.7$ 	& $6807.6$ 	& $7387.4$ 	\tabularnewline
-- 			& $(45,  0)$  	& 	$-$ 		& $6872.0$ 	& $6789.0$ 	& $7346.2$ 	\tabularnewline
-- 			& $(45, 45)$  	& 	$-$ 		& $7154.7$ 	& $6808.5$ 	& $7366.1$ 	\tabularnewline
-torus 		& $( 0,  0)$  	& 	$11843.5$	& $11417.3$ 	& $11792.9$ 	& $12785.4$ 	\tabularnewline
--			& $(45,  0)$  	& 	$-$    	& $11809.3$ 	& $11766.8$ 	& $12844.4$ 	\tabularnewline
-- 			& $(45, 45)$ 	& 	$-$ 	 	& $12086.0$ 	& $11822.4$ 	& $12798.4$ 	\tabularnewline
+Shape      & $(\varphi,\theta)$ & $S$            & $S_{crofton^3}$     & $S_{crofton^{13}}$  & $S_{VTKmarch}$       & $S_{ITKiso}$  \tabularnewline
+\hline 
+\hline 
+ball       & $n.a.$             &     $11309.8$  & $11312.0\ (+0.0\%)$ & $11306.6\ (-0.0\%)$ & $12298.2\ (+8.7\%)$  & $12299.5\ (+8.8\%)$ \tabularnewline 
+prolate    & $( 0,  0)$         &     $3082.9$   & $2938.7\ (-4.7\%)$  & $3084.6\ (+0.1\%)$  & $3354.1\ (+8.8\%)$   & $3354.9\ (+8.8\%)$  \tabularnewline
+-          & $(45,  0)$         &     $-$        & $3137.3\ (+1.8\%)$  & $3086.3\ (+0.1\%)$  & $3353.9\ (+8.8\%)$   & $3354.7\ (+8.8\%)$  \tabularnewline
+-          & $(45, 45)$         &     $-$        & $3150.6\ (+2.2\%)$  & $3083.3\ (+0.0\%)$  & $3351.5\ (+8.7\%)$   & $3352.2\ (+8.7\%)$  \tabularnewline
+oblate     & $( 0,  0)$         &     $6856.8$   & $6290.7\ (-8.3\%)$  & $6807.6\ (-0.7\%)$  & $7410.2\ (+8.1\%)$   & $7411.3\ (+8.1\%)$  \tabularnewline
+-          & $(45,  0)$         &     $-$        & $6872.0\ (+0.2\%)$  & $6789.0\ (-1.0\%)$  & $7369.7\ (+7.5\%)$   & $7370.7\ (+7.5\%)$  \tabularnewline
+-          & $(45, 45)$         &     $-$        & $7154.7\ (+4.3\%)$  & $6808.5\ (-0.7\%)$  & $7385.9\ (+7.7\%)$   & $7386.7\ (+7.7\%)$  \tabularnewline
+torus      & $( 0,  0)$         &     $11843.5$  & $11417.3\ (-3.6\%)$ & $11792.9\ (-0.4\%)$ & $12826.6\ (+8.3\%)$  & $12828.7\ (+8.3\%)$ \tabularnewline
+-          & $(45,  0)$         &     $-$        & $11809.3\ (-0.3\%)$ & $11766.8\ (-0.6\%)$ & $12787.4\ (+8.0\%)$  & $12789.3\ (+8.0\%)$ \tabularnewline
+-          & $(45, 45)$         &     $-$        & $12086.0\ (+1.9\%)$ & $11822.4\ (-0.2\%)$ & $12849.1\ (+8.5\%)$  & $12851.1\ (+8.5\%)$ \tabularnewline
+\hline 
 \end{tabular}\end{center}
@@
 
-% \section{Timing}
-% [c'est rapide]
+\section{Timing}
+
+The performance of the Crofton based method as well as several other methods available in ITK or VTK
+are shown in Table~\ref{tab:perf}. The timings were obtained using the {\tt
+./perf3D ../images/ball450.nrrd} command on an Intel(R) Xeon(R) CPU X5570 processor at 2.93GHz
+with 8192Kb cache, 24Gb of RAM running Ubuntu linux 11.04 64 bits with gcc 4.5.2.
+The input image have a size of $1000 \times 1000 \times 1000$ and contains a single binary object:
+a ball of radius 450.
+
+%./marchingCubes ../images/ball450.nrrd
+% Tcroft    Tmarch    Tmarch2    Timesh    speedupMarch    speedupMarch2    speedupIMesh    Scroft    Smarch    Smarch2    Simesh    %march    %march2    %imesh    fname    
+% 1.80882    26.6482    15.3972    29.1124    14.7324    8.51231    16.0947    2.32339e+06    2.33467e+06    2.52777e+06    2.52779e+06    0.483298    8.08555    8.08621    ../images/ball450.nrrd
+
+\begin{table}[!htb]
+\begin{center}\begin{tabular}{cc}
+Method                                   & Timing \\
+ \hline 
+ \hline
+Proposed method                          & 1.80882 \\
+VTK marching cubes                       & 15.3972 \\
+VTK marching cubes + gaussian filtering  & 26.6482 \\
+ITK BinaryMask3DMeshSource               & 29.1124 \\
+ \hline
+\end{tabular}\end{center}
+\caption{ \label{tab:perf} }
+\end{table}
 

### 2012-02-17 article/Article.tex
fix Peq formula

@@
 roundness             & = & \frac{P_{eq}}{P} \\
-P_{eq}                & = & \frac{d \cdot V_{eq}}{R_{eq}} \\
+P_{eq}                & = & \frac{d \cdot V}{R_{eq}} \\
 R_{eq}                & = & \sqrt[\displaystyle d]{\frac{V\cdot\Gamma(\frac{d+1}{2})}{\pi^\frac{d}{2}}} \\

### 2012-02-20 article/Article.tex
typos dans le doc

@@
 
+% To write tables with a font smaller than text
 \let\oldtabular=\tabular
@@
 whose true perimeter or surface area is known. 
-
 Planar test shapes include disks, rings obtained by the difference of several disks,
-trefoil shape, ellipses with $a=XX$ and $b=XX$, and rectangle with length $XX$ and width $XX$.
-
-Test shapes for 3D measurements include
-balls with various radii, hollow ball, prolate and oblate ellipsoids, and cuboids. 
-For each type of shape, several orientations were considered.
+trefoil shape, ellipses and rectangles with various sizes aspect ratios.
+Test shapes for 3D measurements include balls with various radii, hollow balls, 
+prolate and oblate ellipsoids, and cuboids. Only significant results are presented here.
 
@@
 \begin{center}
-\includegraphics[width=12cm]{images/discreteDisks2}
+\includegraphics[width=10cm]{images/discreteDisks2}
 \end{center}
@@
 
-\subsection{Perimeter of squares with various orientations}
+\subsection{Perimeter of Ellipses with various orientations}
 
-Figure \ref{fig:MeasureSquareOriented} shows measures obtained on discretized squares (side 30 pixels)
-with various orientations. All methods oscillate around the expected value. 
-Using Crofton method with only 2 directions can produce greater errors on the measure than with 4 
+Figure \ref{fig:MeasureSquareOriented} shows measures obtained on discretized ellipses 
+(semi axis lengths equal to 30 and 10) with various orientations. 
+Using Crofton method with only 2 directions produces greater errors on the measure than with 4 
 directions. However, in both cases, the average over the orientations of the mean values is very close 
@@
 
-% Example of reconstruction of a discrete squares with various orientations
+% Example of reconstruction of a discrete ellipses with various orientations
 \begin{figure}[!htb]
@@
 
-% TODO mesure pour des formes 3D
-% able
+The perimeter by counting boundary pixels, for example by using Matlab's "regionprops" methods, 
+leads to a systematic bias: whatever the orientation, measure is different from the actual perimeter.
+Note that the difference does not decrease with the resolution.
 
@@
 
-Comarison of surface area measured on several discrete reconstructed shapes is given in
+Comparison of surface area measured on several discrete reconstructed shapes is given in
 Table \ref{tab:CompareSurfaceArea}.
+Test shapes comprise a ball with radius 30 voxels, 
+prolate ellipsoids (semi axis lengths equal to 30, 10 and 10 voxels), 
+oblate ellipsoids (semi axis lengths equal to 10, 10 and 30 voxels), 
+and torus (outer and inner radius equal to 50 and 10 voxels).
+
 As for the planar case, the measure is closer to the actual value when using more
@@
 1 percent for 13 directions.
-
-When using isosurface reconstruction, the surface area is systematically over estimated.
-The difference is around 8 percent in average.
 
@@
 
+When using Crofton method, the measured perimeter is oscillating around the actual value.
+If 3 directions are used, relative error is usually less than 5 percents.
+Errors are smaller when 13 directions are used (maximum relative error equal to 1 percent).
+
+When using isosurface reconstruction, the surface area is systematically over estimated.
+The difference is around 8 percent in average.
+
 \section{Timing}
@@
 \end{tabular}\end{center}
-\caption{ \label{tab:perf} }
+\caption{ \label{tab:perf} 
+Timing results for measuring the surface area of a ball with radius 450 voxels in a 
+$1000\times 1000 \times 1000$ voxels image.
+}
 \end{table}
-
-\section{Application}
-
-\subsection{ICOPAN}
-
-Methods were applied on images acquired within the ICOPAN project \cite{Andrey2010}.
 

### 2012-02-21 article/Article.tex
ajoute baratin et graphes sur la rondeur, corrections diverses

@@
 
-The roundness is defined as the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
+Roundness is a commom parameter used for comparing shapes independently of their size.
+It equals the square root of the "shape factor", or "isoperimetric deficit".
+It should be noted that although widely used, these parameters present many drawbacks: 
+high sensitivity to perimeter measure, 
+value possibly out of the expected bounds, similar values for different shapes\cite{Pirard2005}...
+
+We defined here the roundness as the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
 The equivalent $d$-perimeter is defined as the perimeter of the $d$-ball with same volume $V$.
@@
 Roundness takes values between $0$ and $1$. 
-It is equal to $1$ for circular or spherical shapes, 
-and decreases for more complicated shapes.
+It is equal to $1$ for circular or spherical shapes, and decreases for more complicated shapes.
+For 3D shapes, equivalent decriptive parameters can be defined using ratio of surface area and volume \cite{Andrey2010}.
 
@@
 
-\subsection{Test shapes}
+%\subsection{Test shapes}
 
@@
 \begin{center}
-\includegraphics[width=10cm]{images/discreteDisks2}
+\includegraphics[width=8cm]{images/discreteDisks2}
 \end{center}
@@
 
-\subsection{Measurements}
+%\subsection{Measurements}
 
@@
 \begin{center}
-\includegraphics[width=10cm]{images/perimeterDisks}
+\includegraphics[width=8cm]{images/perimeterDisks}
 \end{center}
@@
 
-\subsection{Perimeter of Ellipses with various orientations}
+\subsection{Perimeter of ellipses with various orientations}
 
-Figure \ref{fig:MeasureSquareOriented} shows measures obtained on discretized ellipses 
+Figure \ref{fig:PerimeterOrientedEllipses} shows measures obtained on discretized ellipses 
 (semi axis lengths equal to 30 and 10) with various orientations. 
@@
 Left: perimeter measures. Right: relative errors (in percent).}
-\label{fig:MeasureSquareOriented}
+\label{fig:PerimeterOrientedEllipses}
 \end{figure}
@@
 Note that the difference does not decrease with the resolution.
+
+\subsection{Roundness of random ellipses}
+
+Figure \ref{fig:RoundnessRandomEllipses} shows the roundness measured on a population of 100 discretized ellipses with 
+semi axis lengths equal to 4 and 2 pixels. The center of the ellipses is chosen at random within the center pixel, and orientation
+is chosen at random between 0 and 180 degrees.
+
+The variations in the perimeter measure cause variation in the roundness mesure. The variability of roundness is less when measured 
+with the Crofton method using 4 directions. Not that in some cases, the computed roundness may be greater than 1, due to the error in perimeter measure.
+
+% Example of reconstruction of a discrete ellipses with various orientations
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[height=5cm]{images/demoEllipses42}
+%\includegraphics[height=5cm]{images/perimEllipses42Boxplot}
+\includegraphics[height=5cm]{images/roundnessEllipses42Boxplot}
+\end{center}
+\caption{Measure of roundness on random ellipses. Left: some examples of test ellipses and their discrete reconstructions. 
+Right: box plots of roundness computed from perimeter measured from different methods.}
+\label{fig:RoundnessRandomEllipses}
+\end{figure}
 

### 2012-02-21 article/Article.tex
minor fixes

@@
 the method is discussed and tested on several known objects. An algorithm based on the run-length encoding of binary objects is presented
-and compared to a brute force approach.
+and compared to other approaches.
 An implementation is provided and integrated in the LabelObject/LabelMap framework contributed earlier by the authors.
@@
 high sensitivity to perimeter measure, 
-value possibly out of the expected bounds, similar values for different shapes\cite{Pirard2005}...
+value possibly out of the expected bounds, similar values for different shapes\cite{Pirard2005}…
 
@@
 
-The variations in the perimeter measure cause variation in the roundness mesure. The variability of roundness is less when measured 
+The variations in the perimeter measure cause variation in the roundness measure. The variability of roundness is less when measured
 with the Crofton method using 4 directions. Not that in some cases, the computed roundness may be greater than 1, due to the error in perimeter measure.
@@
  \hline
-Proposed method                          & 1.80882 \\
+Proposed method                          & 1.80882 s \\
 VTK marching cubes                       & 15.3972  \\
@@
 \caption{ \label{tab:perf} 
-Timing results for measuring the surface area of a ball with radius 450 voxels in a 
+Excecution times in seconds for measuring the surface area of a ball with radius 450 voxels in a
 $1000\times 1000 \times 1000$ voxels image.
